---
output: git_document
---


```{r, echo = FALSE}
knitr::opts_chunk$set(fig.path = "Figures/EDA/EDA-")
```

## Climate Change Analysis

### Set Preliminaries

```{r}
library(ggplot2)
data_loc="~/Projects/Data/Climate-Change-Analysis"
```

### Load in Data

```{r}
city_data <- read.csv(paste(data_loc, "/GlobalLandTemperaturesByCity.csv", sep = ""))
city_data$dt <- as.Date(city_data$dt, "%Y-%m-%d")
head(city_data)
```

### EDA

There are records for several thousand cities here, each with regular monthly average temperature and uncertainty readings. It will be easier to focus on one city to begin with.

```{r}
minsk_data_init <- city_data[city_data[,"City"]=="Minsk",]
head(minsk_data_init)
```

There are already visible entries with missing data

```{r}
minsk_data_missing = minsk_data_init[is.na(minsk_data_init[,"AverageTemperature"]) | is.na(minsk_data_init[,"AverageTemperatureUncertainty"]),]
dim(minsk_data_missing)
```

Luckily, however, only 73 entries contain missing temperature data. This should be
a small enough proportion of the entries to continue our analysis without needing to
deal with data cleaning.

```{r, echo=FALSE}
print("Missing Dates:")
minsk_data_missing$dt
```
We can see here that the data is complete from around 1753 onwards, with the exception of one month in 2013. The easiest way around this is to remove all entries from before 1753, and ignore the single instance in 2013.

```{r}
minsk_data = minsk_data_init[format(minsk_data_init$dt, format="%Y")>=1753,]
minsk_data = minsk_data[!(is.na(minsk_data[,"AverageTemperature"]) | is.na(minsk_data[,"AverageTemperatureUncertainty"])),]
summary(minsk_data)
```


We can create some plots to observe initial trends in the data

```{r}
plot(minsk_data[, c("dt","AverageTemperature")],type="l",main="Average Temperature in Minsk")
plot(minsk_data[, c("dt","AverageTemperatureUncertainty")],type="l",main="Uncertainty of Average Temperature Reading in Minsk")
```
The average temperature reading data follows a regular variation, as would be
expected throughtout the year. An appropriate next step may be to isolate readings
for each month of the year or seasonally, in order to better see trends across time.

The uncertainty data tells us that early readings for temperature are wildly more
uncertain than those taken in the last century or so.

```{r}
seasonal_months <- c("Jan", "Apr", "Jul", "Oct")
minsk_data_snl <- list()
for (mon in seasonal_months){
    minsk_data_snl[[mon]] <- minsk_data[format(minsk_data$dt, format="%b") == mon,]
}
```
Plot seasonal data, which will make the data easier to model and test:

```{r}
seasonal_plot <- ggplot()
colours = c("#000080","#d40202","#d9d400", "#008000")
for (i in 1:4){
    seasonal_plot <- seasonal_plot + geom_line(data = minsk_data_snl[[i]], aes(x = dt, y = AverageTemperature), color = colours[i], alpha = 0.7)
}
seasonal_plot + theme_minimal()
```

<p>There are no visible trends in the data, as yearly variance causes interference.
However, it is possible we can statistically test for a significant increase over
time, which would indicate provable evidence of global warming.</p>


### Statistical Testing


<p>I want to define an appropriate test for each monthly dataset. There are two ways I could do this: 1) By converting the absolute temperature data into yearly differences, and testing if overall these have a significantly positive mean; 2) By defining a suitable year range (e.g. 1750-1900) as a "control" period, and testing if recent measurements do not fit the control distribution. The first option will be slightly easier to perform, so I will pursue that one initially.</p>

<p>I will first try a basic test, asking if the yearly difference sequence has a mean
greater than zero, which would prove increase over time. I will be using a significance
level of a=0.05. I will try this using each month in turn, thus testing 12 times. I will display my working for January, and extend this to the others.</p>

First, obtain montly data and then convert the absolute data into yearly differences:

```{r}
months <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
minsk_data_monthly <- list()
minsk_data_month_diff <- list()
for (mon in months){
  minsk_data_monthly[[mon]] <- minsk_data[format(minsk_data$dt, format="%b") == mon,]
}
```

```{r}
for (mon in months){
  minsk_data_monthly_diff[[mon]] <- data.frame(minsk_data_monthly[[mon]]$dt[-1])
  names(minsk_data_monthly_diff[[mon]]) = c("Date")
  
  minsk_data_monthly_diff[[mon]]["Difference"] <- minsk_data_monthly[[mon]]$AverageTemperature[-1] - minsk_data_monthly[[mon]]$AverageTemperature[-(dim(minsk_data_monthly[[mon]])[1])]
  
  minsk_data_monthly_diff[[mon]]["DifferenceUncertainty"] <- minsk_data_monthly[[mon]]$AverageTemperatureUncertainty[-1] + minsk_data_monthly[[mon]]$AverageTemperatureUncertainty[-(dim(minsk_data_monthly[[mon]])[1])]
```
