---
output: git_document
---


```{r, echo = FALSE}
knitr::opts_chunk$set(fig.path = "Figures/EDA/EDA-")
```

## Climate Change Analysis

### Set Preliminaries

```{r}
library(ggplot2)
data_loc="~/Projects/Data/Climate-Change-Analysis"
```

### Load in Data

```{r}
city_data <- read.csv(paste(data_loc, "/GlobalLandTemperaturesByCity.csv", sep = ""))
city_data$dt <- as.Date(city_data$dt, "%Y-%m-%d")
head(city_data)
```

### EDA

There are records for several thousand cities here, each with regular monthly average temperature and uncertainty readings. It will be easier to focus on one city to begin with.

```{r}
minsk_data_init <- city_data[city_data[,"City"]=="Minsk",]
head(minsk_data_init)
```

There are already visible entries with missing data

```{r}
minsk_data_missing = minsk_data_init[is.na(minsk_data_init[,"AverageTemperature"]) | is.na(minsk_data_init[,"AverageTemperatureUncertainty"]),]
dim(minsk_data_missing)
```

Luckily, however, only 73 entries contain missing temperature data. This should be
a small enough proportion of the entries to continue our analysis without needing to
deal with data cleaning.

```{r}
minsk_data = minsk_data_init[!(is.na(minsk_data_init[,"AverageTemperature"]) | is.na(minsk_data_init[,"AverageTemperatureUncertainty"])),]
summary(minsk_data)
```


We can create some plots to observe initial trends in the data

```{r}
#dev.new()
plot(minsk_data[, c("dt","AverageTemperature")],type="l",main="Average Temperature in Minsk")
#dev.new()
plot(minsk_data[, c("dt","AverageTemperatureUncertainty")],type="l",main="Uncertainty of Average Temperature Reading in Minsk")
```
The average temperature reading data follows a regular variation, as would be
expected throughtout the year. An appropriate next step may be to isolate readings
for each month of the year or seasonally, in order to better see trends across time.

The uncertainty data tells us that early readings for temperature are wildly more
uncertain than those taken in the last century or so.

```{r}
seasonal_months <- c("Jan", "Apr", "Jul", "Oct")
minsk_data_snl <- list()
for (mon in seasonal_months){
    minsk_data_snl[[mon]] <- minsk_data[format(minsk_data$dt, format="%b") == mon,]
}
```
Plot seasonal data, which will make the data easier to model and test:

```{r}
#dev.new()
seasonal_plot <- ggplot()
colours = c("#000080","#d40202","#d9d400", "#008000")
for (i in 1:4){
    seasonal_plot <- seasonal_plot + geom_line(data = minsk_data_snl[[i]], aes(x = dt, y = AverageTemperature), color = colours[i], alpha = 0.7)
}
seasonal_plot + theme_minimal()
```

There are no visible trends in the data, as yearly variance causes interference.
However, it is possible we can statistically test for a significant increase over
time, which would indicate provable evidence of global warming.

I will first try a basic test, asking if the yearly difference sequence has a mean
greater than zero, which would prove increase over time. I will be using a significance
level of a=0.05
